---
title: "Matrix Formulation & SEM Softwares"
author: "Subhajit Chattopadhyay"
date: "11/6/2020"
output: html_document
---
<link href = "E:/Mine/Books/IT/R-programming/Outline - R Programming/WD/SEM/SEM.css">

<h3>Matrix Formulation - McArdle and McDonald </h3>

<p>There are such procedures, which allow the expression of the equations of a path diagram as the product of several matrices.</p>

<p> Not only does such an approach allow one to turn a path diagram into a set of path equations with less risk of error, but in fact one need not explicitly write down the path equations at all—one can carry out
the calculation of implied correlations directly via operations on the matrices.</p>

<p> Following methods are popular - 

<ul>

<li>McArdle and McDonald (1984)</li>

<li>Benter-Weeks(1980)</li>

<li>McArdle(1980)</li>

<li>McDonald(1978)</li>

</ul>


</p>

<img src = "E:/Mine/Books/IT/R-programming/Outline - R Programming/WD/SEM/Presentation1/Slide21.png">

From the above diagramme following equations can be written - 
$$C = c * B + a * A + e$$
$$D = a * A + d$$
Now the above equaiton can also be written in matrix form in the folloiwng manner - 

$$A = 0*A + 0*B + 0*C + 0*D$$
$$B = 0*A + 0*B + 0*C + B$$
$$C = a * A + c * B + 0 * C + 0 * D + e$$
$$D = a * A + 0 * B + 0 * C + 0 * D + d$$

<img src= "E:/Mine/Books/IT/R-programming/Outline - R Programming/WD/SEM/Presentation1/Slide22.png">

Now, if the vector of A, B C, D is considered as **v** , the coefficients are considered as **A** and the vector D, B , e and d is considered as **u**, then the above matrix can be written as 

$$v = A * v + u$$

$$v * (I - A) = u$$
$$v = (I - A)^{-1}  u$$
Now the covariance / correlation can be expressed as 

$$\frac{vv^{\prime}}{n} = \frac{(I - A)^{-1} u u{^\prime}(I -A)^{-1{^\prime}}}{n}$$
Now , $u * u^{\prime}$ is another covariance matrix , which can be donated as $S$ , so 

$$C = \frac{v*v{^\prime}}{n} = \frac{(I - A)^{-1} * S * (I -A)^{-1{^\prime}}}{n}$$



McArdle and McDonald (1984) defines the structure of the equation in the following manner -

<ul>

<li>A (for symmetric relations) contains paths.</li>

<li>S (for asymmetric relations) contains correlations (or covariances) and residual variances.</li>

<li>F (for filter matrix) selects out the observed variables from the total set of variables.</li>

<img src = "E:/Mine/Books/IT/R-programming/Outline - R Programming/WD/SEM/Presentation1/Slide22.png">

$$C = F * (I - A)^{-1} * S * (I -A)^{-1{^\prime}} * F{^\prime}$$

where , there are "t" variables out of which "m" are measured -

$A$ is a  (t * t) matrix

$S$ is a (t * t) matrix

$F$ is a (m * t) matrix


<h3>Benter-Weeks Model</h3>

$$\eta = \beta * \eta + \gamma * \xi$$
where, there are "q" no of  DVs and "r" no of IVs are there.

<ul>

<li>$\eta$ is a q * q matrix </li>

<li>$\beta$ is a q * q  matrix </li>

<li>$\gamma$ is a q * r matrix </li>

<li>$\xi$ is a r * 1 matrix </li>

<\ul>

<img src= "E:/Mine/Books/IT/R-programming/Outline - R Programming/WD/SEM/Presentation1/Slide7.png">

<p>In the example, SKISAT, SNOWSAT, FOODSAT, NUMYRS, and DAYSKI are all DVs because they all have at least one line with a single-headed arrow pointing to them. Notice that SKISAT is a latent variable and also a dependent variable.</p>

<p>



In the above model , r = 7 and q = 5

If the above above diagramme is represented in terms of the Benter-Weeks model, it would like 


<img src= "E:/Mine/Books/IT/R-programming/Outline - R Programming/WD/SEM/Presentation1/Slide24.png">


<h3>Computer Programmes - SEM</h3>

<ul>

<li>AMOS</li>

<li>LISREL</li>

<li>CALIS</li>

<li>Lavaan</li>

<li>sem</li>

<li>Mx</li>

<li>OpenMx</li>

<li>Mplus</li>

<li>SEPATH</li>

</ul>


<h3>Introduction to Lavaan</h3>

<ul>
<li>Defining the model 

<ul>
<li> Specifying the relationship among measured variables</li>

<li> Specifying the relationship among latent variables</li>

<li> Specifying the relationship among the latent and measured variables </li>

</ul>



<li>Looking the model fit</li>

<li>Visual representation of the model</li>

</ul>


<p>Using Lavvan for FA - Example - 1 </p>
```{r message=FALSE}
library(lavaan)
library(semPlot)
library(psych)
library(psychTools)

test1 <- lavaan::HolzingerSwineford1939

# The data consists of mental ability test scores of e children from two different schools Pasteur and Grant-White.

str(test1)

View(test1)

summary(test1)

test2 <- test1[,-c(1,5,6)]

test.cor <- round(cor(test2),2)

test.cor

library(corrplot)

corrplot(test.cor)

test3 <- test1[,-c(1:6)] # removing all the demographic variables

test.cor1 <- round(cor(test3),2)

test.cor1

cor.plot(test.cor1)

fa1 <- factanal(~test3$x1 + test3$x2 + test3$x3 + test3$x4 + test3$x5 + test3$x4 + test3$x5 + test3$x6 , factors = 3)

fa1


# FA with Lavaan #

model1 <- 'fact1 =~ x1 + x2 + x3
           fact2 =~ x4 + x5 + x6
           fact3 =~ x7 + x8 + x9'


fit <- cfa(model = model1, data = test3)

fit

summary(fit)

semPaths(fit)


```


<h3>SEM wih Lavaan</h3> 

```{r message=FALSE, warning=FALSE}
library(lavaan)
library(semPlot)
library(psych)
library(psychTools)

test4 <- lavaan::PoliticalDemocracy

str(test4)

# y1 - Expert ratings of the freedom of the press in 1960

# y2 - The freedom of political opposition in 1960

# y3 - The fairness of elections in 1960

# y4 - The effectiveness of the elected legislature in 1960

# y5 - Expert ratings of the freedom of the press in 1965

# y6 - The freedom of political opposition in 1965

# y7 - The fairness of elections in 1965

# y8 - The effectiveness of the elected legislature in 1965

# x1 - The gross national product (GNP) per capita in 1960

# x2 - The inanimate energy consumption per capita in 1960

# x3 - The percentage of the labor force in industry in 1960



model2 <- '

# factors 

           eco60 =~ x1 + x2 + x3
           
           pol60 =~ y1 + y2 + y3 + y4
           
           pol65 =~ y5 + y6 + y7 + y8

# regressions

          pol60 ~ eco60 
        
          pol65 ~ eco60 + pol60 '

model2

fit1 <- sem(model= model2, data = test4)

semPaths(fit1)
summary(fit1)

```


```{r message=FALSE, warning=FALSE}

model3 <- '
  # measurement model
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8
  # regressions
    dem60 ~ ind60
    dem65 ~ ind60 + dem60
  # residual correlations
    y1 ~~ y5
    y2 ~~ y4 + y6
    y3 ~~ y7
    y4 ~~ y8
    y6 ~~ y8'


fit2 <- sem(model = model3, data = test4)

semPaths(fit2)

summary(fit2)



```


